A Bayesian ranking-statistic for binary black holes mergers in the second LIGO observing run

### 2021-05-28 main.tex
Updates from Overleaf
add todos

% math keywords 
\newcommand{\fancytext}[1]{{\relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\mathcmd}[1]{{\sc \relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\bcr}{\mathcmd{\rho_\text{BCR}}}
\newcommand{\pycbcstat}{\mathcmd{\rho_\text{PC}}}
\newcommand{\snr}{\mathcmd{\rho}}
\newcommand{\psd}{\mathcmd{P(f)}}
\newcommand{\msun}{\mathcmd{\text{M}_\odot}}
\newcommand{\mtot}{\mathcmd{M}}
\newcommand{\parameters}{\mathcmd{\vec{\theta}}}
\newcommand{\prior}{\mathcmd{\pi(\parameters)}}
\newcommand{\template}{\mathcmd{\mu(\parameters)}}
\newcommand{\fap}{\mathcmd{\text{FAP}}}

@@

\title{A search for high-mass black holes in data from the second LIGO observing run using a Bayesian ranking-statistic}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detectors are plagued with short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes, making their detection difficult. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect high-mass binary black hole mergers (with a total mass $>55$~\msun). We apply this technique to the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some stellar mass binary black holes unreported in GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the `intermediate-mass' range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties due to the observational techniques utilized~\cite{Greene:2020:ARA&A}.

Compact binaries coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). Ground-based gravitational wave detectors can probe the lower-end $10^{2} - 10^{3}\ \msun$ of the intermediate mass range. 
 As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$, systems with large $\text{M}_{T}$ have very low merger frequencies $f < 100\ \text{Hz}$. Hence, ground based gravitational wave detectors ($\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($\sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}.

Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detector data steams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories are plagued with non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Some glitches, similar to signals from high total mass mergers, last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from high total mass mergers~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be removed by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to remove all the glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to account for more glitches while searching for high total mass CBC gravitational-waves is by utilizing a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. In this paper, we utilize a Bayesian method, called the Bayesian Coherence Odds-Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational-wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. The \bcr used in this study is a bootstrap-Bayesian odds computed using Bayesian evidences that describe the explicit probability of data under the hypothesis that the data contains coherent signals vs incoherent glitches, and empirically calculated prior-odds that describe how likely each hypotheses is. \rs{Need to add another couple of sentences describing what is unique about this work. For instance, we study both foreground \textit{and} background events (which is the first time anyone's done a large-scale Bayesian analysis on background data). Suggest using this format: ``We highlight a number of innovations unique to this work. First...''} The high-mass candidate \bcr values are then used to calculate the probability that the candidate is of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for those events (GW170304, GW170727, GW170817A), and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$.

@@

The standard framework to identify CBC gravitational-wave signals in data is by quantifying the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent signal versus just noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal vs. instrumental glitches or noise~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies and populations~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian ``evidence'': a marginalized likelihood. The evidence is the correct likelihood of having obtained the data given a hypothesis. However, due to the high computational cost to calculate the evidence, it is not explicitly used in current matched filter searches. Here, we explore a middle-ground hybrid between frequentist matched filtered searches and Bayesian odds. We compute Bayesian evidences for foreground and background data under the assumption that they either contain coherent gravitational-wave signals, noise, or glitches. However, instead of computing true Bayesian odds, we use the evidences to form a bootstrapped distribution for the odds of foreground and background events to form a frequentist ranking statistic.

@@

Bayes theorem states that the posterior probability distribution $p(\parameters|d,\mathcal{H})$ for data $d$ and a vector of parameters \parameters that describe a model which quantifies a hypothesis $\mathcal{H}$, is given by
\begin{equation}
p(\parameters|d,\mathcal{H}) = \frac{\mathcal{L}(d|\parameters, \mathcal{H}) \ \pi(\parameters | \mathcal{H})}{\mathcal{Z}(d|\mathcal{H})}\ , 
\end{equation}
where $\mathcal{L}(d|\parameters, \mathcal{H})$ is the likelihood of the data given the parameters \parameters and the hypothesis, $\pi(\parameters | \mathcal{H})$ is the prior probability of the parameters, and finally,
\begin{equation}
    \mathcal{Z}(d|\mathcal{H}) = \int\limits_{\parameters} \ \mathcal{L}(d|\parameters,\mathcal{H}) \ \pi(\parameters | \mathcal{H}) \ d\parameters\ ,
\end{equation} is the likelihood after marginalizing over the parameters \parameters.  To compare two hypotheses $\mathcal{H}_A$ and $\mathcal{H}_B$ with the Bayes theorem one can calculate an odds-ratio
\begin{equation}
    \mathcal{O}^A_B = \frac{\mathcal{Z}^A\ \pi^A}{\mathcal{Z}^B\ \pi^B}\ ,
\end{equation}
where  $\{\pi^A, \pi^B\}$ are the prior-odds for each hypothesis and $\{\mathcal{Z}^A, \mathcal{Z}^B\}$ are shorthand for the evidences $\{\mathcal{Z}(d|\mathcal{H}_A), \mathcal{Z}(d|\mathcal{H}_B)\}$. The odds-ratio can quantify which of the two hypotheses is more likely. For example, if $\mathcal{O}^A_B >> 1$, then the odds are in favor of the $\mathcal{H}_A$ hypotheses.

The \bcr is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\mathcal{O}_{\mathrm{BCR}} = \frac{P^S Z^S}{\prod\limits^D_{i=1} \ [P^G Z^G_i + (1-P^G)Z^N_i]}\ ,
\end{equation}

where $P^S$ and $P^G$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $P^S=P(\mathcal{H}_S)/P(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $P^G=P(\mathcal{H}_G| \mathcal{H}_I)$, the probability of obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

\begin{equation}
\label{eq:bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ ,
\end{equation}

\avi{Ranking statistic \bcr is like an estimator of what $\mathcal{O}_{\mathrm{BCR}}$ is}

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the correct prior-odds are known, the \bcr is a Bayesian odds-ratio. As an odds-ratio, the \bcr is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

When using the \bcr as a detection statistic, the physical interpretation of the prior-odds is lost. Hence, the prior-odds are empirically tuned to maximize the separation between the \bcr distribution of the background (expected to favor the $\mathcal{H}_I$ hypothesis) and the \bcr distribution of artificially manufactured simulated signals (expected to favor the $\mathcal{H}_S$ hypothesis). Increasing the separation between the two distributions can improve ability of the \bcr to discriminate candidate events as coherent signals or incoherent instrumental features. The tuning process is described in detail in Appendix~\ref{sec:tuning-prior-odds}.

\subsection{Estimation of astrophysical signal probability}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors, \fap, for the candidate \bcr-value can quantify the significance. The \fap is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal.

To calculate the \fap, each candidate \bcr is considered a single statistical trial that can occur at a fixed false alarm probability $f$, where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}
The false alarm probability with trials \fap that the \bcr measurement occurs at least once for $N$ trials $(N > 0)$, where $N$ is the number of candidate triggers is
\begin{equation}
    \fap = 1 - (1-f)^N \ .
\end{equation}

@@

In addition to finding candidate events, \pycbc also identified several million background triggers for each time-frame, by searching background data manufactured by time-sliding data within that time-frame. The background triggers help quantify the candidate events' significance for the respective time-frames. Finally, to test the search's sensitivity, \pycbc produced and searched for thousands of simulated signals.

>>>>

% math keywords 
\newcommand{\fancytext}[1]{{\relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\mathcmd}[1]{{\sc \relax\ifmmode#1\else $#1$\fi}\xspace}
\newcommand{\bcr}{\mathcmd{\rho_\text{BCR}}}
\newcommand{\bcrodds}{\mathcmd{\mathcal{O}_{\mathrm{BCR}}}}
\newcommand{\pycbcstat}{\mathcmd{\rho_\text{PC}}}
\newcommand{\snr}{\mathcmd{\rho}}
\newcommand{\psd}{\mathcmd{P(f)}}
\newcommand{\msun}{\mathcmd{\text{M}_\odot}}
\newcommand{\mtot}{\mathcmd{M}}
\newcommand{\parameters}{\mathcmd{\vec{\theta}}}
\newcommand{\prior}{\mathcmd{\pi(\parameters)}}
\newcommand{\template}{\mathcmd{\mu(\parameters)}}
\newcommand{\fap}{\mathcmd{\text{FAP}}}

@@

\title{A Bayesian ranking-statistic for binary black holes mergers in the second LIGO observing run}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detector data contain numerous short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes, making intermediate-mass merger detection challenging. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect binary black hole mergers with a total mass $>55$~\msun. \avi{Say why it improves on prev techniques}. We apply this technique to candidate events with total masses $>55$~\msun during LIGO's second observing run to search for previously unresolved gravitational-wave signals from this heavy category of binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some stellar-mass binary black holes unreported in GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the `intermediate-mass' range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties intrinsic to the observational techniques~\cite{Greene:2020:ARA&A}.

Compact binaries coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$,  ground based gravitational wave detectors ($f\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($f \sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}. Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detector data streams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories data-streams have numerous non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Like signals from intermediate mass mergers, some glitches last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from black hole mergers with total masses $>55$~\msun ~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be removed by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to remove all the glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to account for more glitches while searching for CBC gravitational-waves is to rely on a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. In this paper, we rank O2's candidate gravitational-wave signals from $55-500\ \msun$ total mass systems using a Bayesian method called the Bayesian Coherence Odds-Ratio \bcr~\cite{BCR1}. \avi{Should explain: whats the advantage of this? Why is it useful for triggers with $M_T>55\msun$?} The \bcr used in this study is a bootstrap-Bayesian odds calculated using the probability of data under the hypothesis that the data contains coherent signals vs. incoherent glitches and empirically calculated prior-odds that describe how likely each hypothesis is. After computing the candidate \bcr values, we calculate the candidate's probability of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.  We highlight several innovations unique to this work. First, we use our Bayesian framework to analyze all candidate signals and background triggers (incoherent signal-free data) with total masses above $> 55 \msun$ in O2. Second, we empirically tune prior-odds for finding a signal or a glitch in O2 using \bcr distributions for background triggers and simulated signals. Finally,  we use the tuned-\bcr background distribution to compute how significant candidate signals are from the background.

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for those events (GW170304, GW170727, GW170817A); and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$ \avi{Need to doublecheck that this wasnt in the PyCBC sub-threshold paper}.

@@

The standard framework to identify CBC gravitational-wave signals in data is to quantify the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. \avi{Holdup: isnt GSTLAL's method a bootstrapped-Bayesian method?} Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent signal versus just noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

\avi{I discuss the benefits of the BCR here...}
Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal vs. instrumental glitches or noise~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies and populations~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian ``evidence'': a marginalized likelihood. The evidence is the correct likelihood of having obtained the data given a hypothesis. However, due to the high computational cost to calculate the evidence, it is not explicitly used in current matched filter searches. Here, we explore a middle-ground hybrid between frequentist matched filtered searches and Bayesian odds. We compute Bayesian evidences for foreground and background data under the assumption that they either contain coherent gravitational-wave signals, noise, or glitches. However, instead of computing true Bayesian odds, we use the evidences to form a bootstrapped distribution for the odds of foreground and background events to form a frequentist ranking statistic.

@@

Bayes' theorem states that the posterior probability distribution $p(\parameters|d,\mathcal{H})$ for data $d$ and a vector of parameters \parameters that describe a model which quantifies a hypothesis $\mathcal{H}$, is given by
\begin{equation}
p(\parameters|d,\mathcal{H}) = \frac{\mathcal{L}(d|\parameters, \mathcal{H}) \ \pi(\parameters | \mathcal{H})}{\mathcal{Z}(d|\mathcal{H})}\ , 
\end{equation}
where $\mathcal{L}(d|\parameters, \mathcal{H})$ is the likelihood of the data given the parameters \parameters and the hypothesis, $\pi(\parameters | \mathcal{H})$ is the prior probability of the parameters, and finally,
\begin{equation}
    \mathcal{Z}(d|\mathcal{H}) = \int\limits_{\parameters} \ \mathcal{L}(d|\parameters,\mathcal{H}) \ \pi(\parameters | \mathcal{H}) \ d\parameters\ ,
\end{equation} is the likelihood after marginalizing over the parameters \parameters.  To compare two hypotheses $\mathcal{H}_A$ and $\mathcal{H}_B$ through Bayes' theorem one can calculate an odds-ratio
\begin{equation}
    \mathcal{O}^A_B = \frac{\mathcal{Z}^A\ \pi^A}{\mathcal{Z}^B\ \pi^B}\ ,
\end{equation}
where  $\{\pi^A, \pi^B\}$ are the prior-odds for each hypothesis and $\{\mathcal{Z}^A, \mathcal{Z}^B\}$ are shorthand for the evidences $\{\mathcal{Z}(d|\mathcal{H}_A), \mathcal{Z}(d|\mathcal{H}_B)\}$. The odds-ratio can quantify which of the two hypotheses is more likely. For example, if $\mathcal{O}^A_B >> 1$, then the odds are in favor of the $\mathcal{H}_A$ hypotheses.

The \bcrodds quantity is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, \bcrodds is given by
\begin{equation}
\label{eq:bcr}
\bcrodds = \frac{\pi^S Z^S}{\prod\limits^D_{i=1} \ [\pi^G Z^G_i + (1-\pi^G)Z^N_i]}\ ,
\end{equation}

where $\pi^S$ and $\pi^G$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $\pi^S=\pi(\mathcal{H}_S)/\pi(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $\pi^G=\pi(\mathcal{H}_G| \mathcal{H}_I)$, the probability of obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the prior-odds represent our true beliefs, the \bcrodds is a Bayesian odds-ratio. As an odds-ratio, the \bcrodds is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, we \textit{tune} values for prior-odds, $\hat{\pi}^S$ and $\hat{\pi}^G$, to estimate \bcrodds with a ranking statistic, \bcr, given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ .
\end{equation}

In the limit where the estimated prior-odds equal the true prior-odds, $\bcr\to\bcrodds$. However, as we are uncertain what the true prior-odds are, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

When using the \bcr as a detection statistic, the prior-odds are empirically tuned to maximize the separation between the \bcr distribution of the background (expected to favor the $\mathcal{H}_I$ hypothesis) and the \bcr distribution of artificially manufactured simulated signals (expected to favor the $\mathcal{H}_S$ hypothesis). Increasing the separation between the two distributions can improve ability of the \bcr to discriminate candidate events as coherent signals or incoherent instrumental features. The tuning process is further described in Appendix~\ref{apdx:tuning-prior-odds}.

\subsection{Estimation of astrophysical signal probability}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant relative to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors \avi{(need to define a trial factor)}, \fap, for the candidate \bcr-value can quantify the significance. The \fap is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal.

To calculate the \fap, each candidate \bcr is considered a single statistical trial that can occur at a fixed false alarm probability $f$, where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}
The false alarm probability with trials \fap that the \bcr measurement occurs at least once for $N$ trials $(N > 0)$, where $N$ is the number of candidate signals, is \avi{(This is very convoluted)}
\begin{equation}
    \fap = 1 - (1-f)^N \ .
\end{equation}

@@

In addition to finding candidate events, \pycbc also identified several million background triggers for each time-frame, by searching background data manufactured by time-sliding data within that time-frame. The background triggers help quantify the candidate events' significance for the respective time-frames. Finally, to test the search's sensitivity, \pycbc produced and searched for thousands of simulated signals. \avi{Do i need to contextualize this choice more, ie talk about what would happen if I chose a narrower range?}
